---
title: "R Notebook"
output: html_notebook
---



```{r}
rm(list = ls())
```

Load and check data

```{r}
# Load packages
library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
```


Now that our packages are loaded, let’s read in and take a peek at the data.

```{r}
train <- read.csv('train.csv', stringsAsFactors = F, encoding = 'UTF-8', na.strings = c('', 'NA', NULL))
test  <- read.csv('test.csv', stringsAsFactors = F, encoding = 'UTF-8', na.strings = c('', 'NA', NULL))
full  <- bind_rows(train, test) # bind training & test data
```

#PassengerId - A unique Id for each passenger. Each Id takes the form gggg_pp where gggg indicates a group the passenger is travelling with and pp is their number within the group. People in a group are often family members, but not always.

#HomePlanet - The planet the passenger departed from, typically their planet of permanent residence.

#CryoSleep - Indicates whether the passenger elected to be put into suspended animation for the duration of the voyage. Passengers in cryosleep are confined to their cabins.

#Cabin - The cabin number where the passenger is staying. Takes the form deck/num/side, where side can be either P for Port or S for Starboard.

#Destination - The planet the passenger will be debarking to.

#Age - The age of the passenger.

#VIP - Whether the passenger has paid for special VIP service during the voyage.

#RoomService, FoodCourt, ShoppingMall, Spa, VRDeck - Amount the passenger has billed at each of the Spaceship Titanic's many luxury amenities.

#Name - The first and last names of the passenger.

#Transported - Whether the passenger was transported to another dimension. This is the target, the column you are trying to predict.

```{r}
# check data
str(full)
```

```{r}
summary(full)
```

#Feature Engineering

#Gonna make a new column with just the surname
```{r}
full$Surname <- gsub('.* ', '', full$Name)
head(full$Surname)
```
#Success!

#PassengerId - A unique Id for each passenger. Each Id takes the form gggg_pp where gggg indicates a group the passenger is travelling with and pp is their number within the group. People in a group are often family members, but not always.

#Gonna make a groupnumber column and a number within group column!
```{r}
full$Group <- as.numeric(gsub('_.*', '', full$PassengerId))
full$Num_in_group <- as.numeric(gsub('.*_', '', full$PassengerId))

head(full[,c('Group', 'Num_in_group')])
```

#Cabin - The cabin number where the passenger is staying. Takes the form deck/num/side, where side can be either P for Port or S for Starboard.

#Gonna make 3 difference columns of the Cabin column.
```{r}
full$Deck <- sub('/.*', '', full$Cabin) #Success
full$Cabin_side <- gsub('(.*/)|(/.*)', '', full$Cabin) #"|" means what i want in the middle
#NVM i accidentaly got side from this


library(stringr)

full$Cabin_num <- str_extract(full$Cabin, "(?<=/).*?(?=/)") #OMG I DID IT
#I GOT IT FROM THE MIDDLE!!
#https://stackoverflow.com/questions/56452306/how-to-extract-the-middle-part-of-a-string-in-a-data-frame-in-r
```


```{r}
full$Cabin_num <- as.numeric(full$Cabin_num)
str(full)
```

#Now i wanna create factor variables (Category variables)

```{r}
# Make variables factors into factors
factor_vars <- c('HomePlanet','CryoSleep','Destination','VIP', 
                 'Transported','Surname','Deck','Cabin_side')

full[factor_vars] <- lapply(full[factor_vars], function(x) as.factor(x))
```


```{r}
str(full)
```

#Gonna make a new column that counts how many of that particular Surname exists in dataset
```{r}
hmm <- full %>%
  group_by(Surname) %>%
  summarise(Totalsur = length(Surname))

tmp <- merge(full, hmm, by = "Surname") #This is a join in SQL

tmp[tmp$Totalsur > 200, 'Totalsur'] <- NA

full <- tmp

remove(tmp) #I DID IT!!
```

#PassengerId - A unique Id for each passenger. Each Id takes the form gggg_pp where gggg indicates a group the passenger is travelling with and pp is their number within the group. People in a group are often family members, but not always.

#Gonna make one "is probably family" variable, if they have the same surname and are in the same group.
```{r}
hmm <- full %>% #First im gonna make a how many is in the group count
  group_by(Group) %>%
  summarise(Totalgrp = length(Group))

tmp <- merge(full, hmm, by = "Group") #Detta är typ en join som jag körde

full <- tmp

remove(tmp)
```

```{r}
hmm <- full %>% #Score!
  group_by(Group, Surname) %>%
  summarise(Probtotfam = length(Surname))

# Create a group_surname variable to be able to join successfully 
full$Group_surname <- paste(full$Group, full$Surname, sep='_')

# Create a group_surname variable to be able to join successfully 
hmm$Group_surname <- paste(hmm$Group, hmm$Surname, sep='_')

hmm <- hmm[,c("Group_surname", "Probtotfam")]

tmp <- merge(full, hmm, by = "Group_surname") #Detta är typ en join som jag kör

full <- tmp

remove(tmp)
remove(hmm)
```

#Taking another look at the data!

```{r}
summary(full)
```

#Creating a how many in this cabin column
```{r}
hmm <- full %>%
  group_by(Cabin) %>%
  summarise(Tot_cabin_ppl = length(Cabin))

tmp <- merge(full, hmm, by="Cabin")

full <- tmp

remove(tmp)
remove(hmm)
```


#Looking through the data, it seems that those that are in the same cabin has the same homeplanet, good for fixing NA values for homeplanet!
```{r}
hmm <- full %>%
  group_by(Cabin) %>%
  summarise(length(HomePlanet), length(Cabin))

uhh <- full[!is.na(full$Cabin) & !is.na(full$HomePlanet),c('Cabin', 'HomePlanet')]

thinking <- merge(hmm, uhh, by='Cabin')
jeez <- thinking[!duplicated(thinking),] #Remove duplicates

jeez <- jeez %>%
  mutate(HomePlanet_2 = HomePlanet)

jeez <- jeez[,c('Cabin', 'HomePlanet_2')]


tmp <- left_join(full, jeez, by = c("Cabin"="Cabin"))#Need to leftjoin

full <- tmp

```


```{r}
remove(tmp)
remove(jeez)
remove(thinking)
remove(hmm)
remove(uhh)
```


```{r}
full[is.na(full$HomePlanet) & !is.na(full$Cabin), 'HomePlanet'] <- 
  full[is.na(full$HomePlanet) & !is.na(full$Cabin), 'HomePlanet_2']
```

#Lets see if it worked!
```{r}
summary(full$HomePlanet) # I think it worked a little bit atleast?
#from 288NA's to 189
```

```{r}
full[is.na(full$HomePlanet) & !is.na(full$Cabin),]
```


```{r}
full[full$Group==8322,]
```
```{r}
#Fixing this NA
full[full$PassengerId =='8322_02', 'HomePlanet'] <- 'Europa'
```

#Checking for more easy NA fixes.
```{r}
full[is.na(full$HomePlanet) & !is.na(full$Cabin) & full$Probtotfam > 1,] #Not sure how to fix this maybe group by group_surname and get that HomePlanet?
```

#Here i am clumping togheter so everyone that is in the same group and has the same surname
#gets the same HomePlanet, this is to fix some NAs
```{r}
hmm <- full %>%
  group_by(Group_surname) %>%
  summarise(length(HomePlanet), length(Group_surname))

uhh <- full[!is.na(full$Group_surname) & !is.na(full$HomePlanet),
            c('Group_surname', 'HomePlanet')]

thinking <- merge(hmm, uhh, by='Group_surname')
jeez <- thinking[!duplicated(thinking),] #Remove duplicates

jeez <- jeez %>%
  mutate(HomePlanet_3 = HomePlanet)

jeez <- jeez[,c('Group_surname', 'HomePlanet_3')]


tmp <- left_join(full, jeez, by = c("Group_surname"="Group_surname"))#Need to leftjoin

full <- tmp
```


```{r}
full[is.na(full$HomePlanet) & !is.na(full$Group_surname), 'HomePlanet'] <- 
  full[is.na(full$HomePlanet) & !is.na(full$Group_surname), 'HomePlanet_3']
```


```{r}
summary(full$HomePlanet) #Got some less NAs
```

```{r}
remove(tmp)
remove(jeez)
remove(thinking)
remove(hmm)
remove(uhh)
```

#Checking if i can see anymore ways to fix homeplanet NAs

```{r}
inspect <- full[is.na(full$HomePlanet),]
#View(inspect) #Doesnt seem like it, could check if there is some connection with Deck and homeplanet
remove(inspect)
```

#Time to plot stuff, to maybe get a feel!
```{r}
ggplot(data = full, aes(x = c(Deck), fill = factor(HomePlanet))) +
  geom_bar() +
  theme_few()
```

#Looks like we can fix some more NAs!!

```{r}
full[full$Deck %in% 'G' & is.na(full$HomePlanet), 'HomePlanet'] <- 'Earth'
#woah you have to use %in% when you want to convert NAs

full[full$Deck %in% 'A' & is.na(full$HomePlanet), 'HomePlanet'] <- 'Europa'
#woah you have to use %in% when you want to convert NAs
full[full$Deck %in% 'B' & is.na(full$HomePlanet), 'HomePlanet'] <- 'Europa'
#woah you have to use %in% when you want to convert NAs
full[full$Deck %in% 'C' & is.na(full$HomePlanet), 'HomePlanet'] <- 'Europa'
#woah you have to use %in% when you want to convert NAs
```

#Lets see if there is some connection with Destination and homeplanet!
```{r}
ggplot(data = full, aes(x = Destination, fill = factor(HomePlanet))) +
  geom_bar() +
  theme_few()
```

#Looks like we can fix some more NAs!!
#We can be pretty sure those who goes to PSO are from Earth

```{r}
full[full$Destination %in% 'PSO J318.5-22' & is.na(full$HomePlanet), 'HomePlanet'] <- 'Earth'
#woah you have to use %in% when you want to convert NAs
```


#Lets see how many NAs we have left now

```{r}
summary(full$HomePlanet)
```

#Lets see if there is any connection with Cabin side and HomePlanet!

```{r}
ggplot(data = full, aes(x = Destination, fill = factor(Cabin_side))) +
  geom_bar() +
  theme_few()
```
#Nope! Taking one more look at the NAs!

```{r}
full[is.na(full$HomePlanet),]
```
#I think we now can assume that those in the same group has the same HomePlanet!

```{r}
full[full$Group == '64',]
```


```{r}
hmm <- full %>%
  group_by(Group) %>%
  summarise(length(HomePlanet), length(Group))

uhh <- full[!is.na(full$Group) & !is.na(full$HomePlanet),c('Group', 'HomePlanet')]

thinking <- merge(hmm, uhh, by='Group')
jeez <- thinking[!duplicated(thinking),] #Remove duplicates

jeez <- jeez %>%
  mutate(HomePlanet_4 = HomePlanet)

jeez <- jeez[,c('Group', 'HomePlanet_4')]


tmp <- left_join(full, jeez, by = c("Group"="Group"))#Need to leftjoin

full <- tmp
```

```{r}
full[is.na(full$HomePlanet) & !is.na(full$Group), 'HomePlanet'] <- 
  full[is.na(full$HomePlanet) & !is.na(full$Group), 'HomePlanet_4']
```

```{r}
remove(tmp)
remove(jeez)
remove(thinking)
remove(hmm)
remove(uhh)
```

```{r}
summary(full$HomePlanet)
```
#Even less NA's!
#Lets leave this for now and take a look at destination NAs!

```{r}
summary(full$Destination)
```

```{r}
#View(full[,c('Group', 'Destination')])
```

#Destination and group doesnt seem to correlate...

#Lets try and plot out if destination has anything to do with homeplanet!
```{r}
ggplot(data = full, aes(x = HomePlanet, fill = factor(Destination))) +
  geom_bar() +
  theme_few()
```
#Hm i think we can assume here that the missing destination values from Mars goes to TRAPPIST-1e
```{r}
full[is.na(full$Destination) & full$HomePlanet %in% 'Mars', 'Destination'] <- 'TRAPPIST-1e'
```


```{r}
summary(full$Destination) #Progress
```

#Fix tot_cabin_ppl, those that miss cabin got high tot_cabin_ppl, time to make those to NA
```{r}
full[full$Tot_cabin_ppl > 200, 'Tot_cabin_ppl'] <- NA
```



#Removing unnecessary columns
```{r}
full$HomePlanet_2 <- NULL
full$HomePlanet_3 <- NULL
full$HomePlanet_4 <- NULL
```

#Im to tired to keep on feature engineering so im gonna fix the rest of the variables with some mice, then im done with it!
#AKA some predicted imputation!

#I’m going to use the mice package for this task

```{r}
# Set a random seed
set.seed(1337)

# Perform mice imputation, excluding certain less-than-useful variables:
mice_mod <- mice(full[, names(full) %in% c('Group_surname','Group','HomePlanet','CryoSleep','Destination','Age','VIP', 'RoomService',
  'FoodCourt', 'ShoppingMall', 'Spa', 'VRDeck', 'Deck', 'Cabin_side', 'Cabin_num')], method='rf')
#method='rf' probably means randomforest!
```


```{r}
# Save the complete output 
mice_output <- complete(mice_mod) #Here it becomes its own dataframe!
```

#Let’s compare the results we get with the original distribution of the variables

```{r}
# Plot age distributions
par(mfrow=c(1,2))
hist(full$Age, freq=F, main='Age: Original Data', 
  col='darkgreen', ylim=c(0,0.04))
hist(mice_output$Age, freq=F, main='Age: MICE Output', 
  col='lightgreen', ylim=c(0,0.04))
```

```{r}
# Plot roomservice distributions
par(mfrow=c(1,2))
hist(full$RoomService, main='RoomService: Original Data', 
  col='darkgreen', breaks = 80, freq=F)
hist(mice_output$RoomService, freq=F, main='RoomService: MICE Output', 
  col='lightgreen', breaks = 80)
```

#Gonna check VIP and cryosleep aswell
```{r}
summary(full$CryoSleep)
```

```{r}
summary(mice_output$CryoSleep)
```
```{r}
summary(full$VIP)
```


```{r}
summary(mice_output$VIP)
```
#Looks good tbh!

```{r}
summary(mice_output)
```

```{r}
summary(full)
```

#What i can see, it looks cool!

#Lets see Deck

```{r}
summary(full$Deck)
```

```{r}
summary(mice_output$Deck)
```

#Everything looks fine! Now lets turn in into the full dataset!
```{r}
full[, names(full) %in% c('Group_surname','Group','HomePlanet','CryoSleep','Destination','Age','VIP', 'RoomService','FoodCourt', 'ShoppingMall', 'Spa', 'VRDeck', 'Deck', 'Cabin_side', 'Cabin_num')] <- mice_output
```


```{r}
#Done
```


```{r}
# First we'll look at the relationship between age & transportation
ggplot(full[!is.na(full$Transported),], aes(Age, fill = factor(Transported))) + #1 till 8693 is the training data
  geom_histogram() + 
  # I also include Cryosleep and VIP to see
  facet_grid(.~CryoSleep) + #Group everything by CryoSleep
  theme_few()
```


```{r}
# First we'll look at the relationship between age & transportation
ggplot(full[!is.na(full$Transported),], aes(Age, fill = factor(Transported))) + 
  geom_histogram() + 
  # I also include Cryosleep and VIP to see
  facet_grid(.~VIP) + #Group everything by VIP
  theme_few()
```
#VIP doesnt seem to matter

#I can see that cryosleep matters and if its a child <10 or a teenager <20

```{r}
# First we'll look at the relationship between age & transportation
ggplot(full[!is.na(full$Transported),], aes(Age, fill = factor(Transported))) + #1 till 8693 is the training data
  geom_histogram() + 
  facet_grid(.~Deck) + #Group everything by Deck
  theme_few()
```
#Deck seems to matter too

```{r}
# Create the column child, and indicate whether child, teenager or adult
full$Child[full$Age < 10] <- 'Child'
full$Child[full$Age < 20 & full$Age >= 10] <- 'Teenager'
full$Child[full$Age >= 20] <- 'Adult'

full$Child  <- factor(full$Child)
```


```{r}
#looking at the relationship between agebracket & transportation
ggplot(full[!is.na(full$Transported),], aes(Child, fill = factor(Transported))) +
  geom_bar() + 
  theme_few()
```
#It's clear its an advantage being a child!

#From earlier tries at prediction i can see that Spa, VRDeck and RoomService had the biggest importance, im gonna try and see if its possible to feature engineer it.
#NOTE: I ended up not using these because they prediction got worse.
#Probably because of over-fitting?


```{r}
ggplot(full[!is.na(full$Transported),], aes(Spa, fill = factor(Transported))) +
  geom_histogram() +
  scale_x_continuous(limits = c(NA, 50)) +
  theme_few()
```
#Looks like i can make a categorical variable that says if the person spent money more than 5 on the spa or not

```{r}
full$Spa_dummy <- NA
full[full$Spa > 5, 'Spa_dummy'] <- 1
full[full$Spa <= 5, 'Spa_dummy'] <- 0
```


```{r}
# Lets now see VRDeck
ggplot(full[!is.na(full$Transported),], aes(VRDeck, fill = factor(Transported))) +
  geom_histogram() +
  scale_x_continuous(limits = c(NA, 10)) +
  theme_few()

```
#Seems to be a pattern here!
#From playing around with the graph, it seems like a person gets punished if they spend more than 5 at the VRDeck.

```{r}
full$VRDeck_dummy <- NA
full[full$VRDeck > 5, 'VRDeck_dummy'] <- 1
full[full$VRDeck <= 5, 'VRDeck_dummy'] <- 0
```

```{r}
# Lets now see RoomService
ggplot(full[!is.na(full$Transported),], aes(RoomService, fill = factor(Transported))) +
  geom_histogram() +
  scale_x_continuous(limits = c(NA, 10)) +
  theme_few()
```
#Here it seems like spending any money at all gets punished!

```{r}
full$RoomService_dummy <- NA
full[full$RoomService > 0, 'RoomService_dummy'] <- 1
full[full$RoomService == 0, 'RoomService_dummy'] <- 0
```

```{r}
summary(full$Probtotfam)
```


```{r}
summary(full$Cabin_num) #Need to fix these in imputation #UPDATE FIXED
```

#Splitting the data back into the original test and training sets.

```{r}
# Split the data back into a train set and a test set
test <- full[is.na(full$Transported),]
train <- full[!is.na(full$Transported),]
```


```{r}
#Splitting the train set into training_train set and training_test set
library(caTools)

split <- sample.split(train$Transported, SplitRatio = 0.8)
training_train <- subset(train, split == TRUE)
training_test <- subset(train, split == FALSE)
```


```{r}
# Setting a random seed
set.seed(1337)

# Building the model
rf_model <- randomForest(factor(Transported) ~ CryoSleep + Deck + Age + HomePlanet +    
                                            RoomService + FoodCourt + ShoppingMall + 
                                            Spa + VRDeck + VIP + Num_in_group +
                                            Cabin_side + Destination + Child + Probtotfam +
                                            Cabin_num + Num_in_group + Totalgrp, 
                                            data = training_train, ntree = 5000, 
                                            nodesize = 8)

```

#Let’s look at relative variable importance by plotting the mean decrease in Gini calculated across all trees.

```{r}
# Getting importance
importance    <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))

# Creating a rank variable based on importance
rankImportance <- varImportance %>%
  mutate(Rank = paste0('#',rank(desc(Importance))))

# Using ggplot2 to visualize the relative importance of variables
ggplot(rankImportance, aes(x = reorder(Variables, Importance), #Reorder makes lowest best or "highest"
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()
```

```{r}
training_prediction <- predict(rf_model, training_test) #Predicting my 
```

```{r}
# Saving the solution to a dataframe with two columns: PassengerId and Survived (prediction)
training_solution <- data.frame(PassengerId = training_test$PassengerId, 
                                Transported = training_prediction)
```


```{r}
training_test_compare <- training_test %>% #This is what i compare my solution, for finetune
  select(c('PassengerId', 'Transported'))

training_test_compare$Rank <- rank(training_test_compare$PassengerId)

training_test_compare <- training_test_compare %>%
  arrange(Rank)
```


```{r}
training_solution$Rank <- rank(training_solution$PassengerId)

training_solution <- training_solution %>%
  arrange(Rank)

rownames(training_solution) <- 1:nrow(training_solution) #Resetting index
```


```{r}

result <- 0
for(i in 1:nrow(training_solution)){
  if(training_solution$Transported[[i]][1] != training_test_compare$Transported[[i]][1]){
    result <- result + 1
  }
}

print(1 - result/nrow(training_solution))
```

#v16 80.1%!


```{r}
# Setting a random seed
set.seed(1337)

# Building the model
rf_model <- randomForest(factor(Transported) ~ CryoSleep + Deck + Age + HomePlanet +    
                                            RoomService + FoodCourt + ShoppingMall + 
                                            Spa + VRDeck + VIP + Num_in_group +
                                            Cabin_side + Destination + Child + Probtotfam +
                                            Cabin_num + Num_in_group + Totalgrp, 
                                            data = train, ntree = 5000, 
                                            nodesize = 8)

#After some experiments, this is my 16th version.
```

#Real and final prediction!
```{r}
# Predict using the test set
prediction <- predict(rf_model, test)

# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(PassengerId = test$PassengerId, Transported = prediction)

# Writing the solution to file
write.csv(solution, file = 'finalprediction.csv', row.names = F)
```

#kaggle score: 0.80313